### 2017-11-10 paper/submit/referee-response.txt
response to ref

We thank the referee for the constructive report and suggested improvements. To
address the referee's concerns, we have redone and rewritten our population
modeling to provide a probabilistic classification of each pair as being
"comoving" or "field." Though our conclusions or results haven't changed as a
result of this change, we agree that it is more robust and more honest to
provide probabilities for each pair rather than perform a hard cut on likelihood
ratio increase. We believe that the re-analysis and re-write of our modeling
section will address many of the broader concerns the referee had with this work
and thus don't reply to every issue specifically. We've also moved away from
using the word "confirmed comoving pairs" and instead say "high-probability."
See especially sections 3 and 4 in the revised draft.

To assist the referee in interpreting the differences, we have also uploaded a
"diff"ed and rendered version of the paper that shows the changes made
(diff.pdf).

Below we address additional, specific concerns raised in the report: lines that
contain questions or comments raised by the referee begin with ">" and our
responses are listed immediately below.

> Minor concerns:
> ===============

> (1) The comparison with LAMOST metallicity and Delta-RV seems important enough
> for the case that it should probably be detailed in its own subsection, rather
> than being placed in the discussion.

We agree, and have moved it into the "Results" section - see Section 4.4

> (2) The description of the data reduction in the appendices is perhaps
> too detailed and too general, especially the 1D spectrum extraction
> and wavelength calibration. If this is standard procedure for
> MDM-MODSPEC reduction, then these details make the paper
> unnecessarily long. Publishing these details might however be useful
> if this is a novel reduction procedure - but this does not seem to be
> the case. The RV measurements on the other hand might be worth
> detailing if the procedure is truly novel, or if it is critical in
> demonstrating the level of RV precision needed for the study.

We aren't aware of a spectroscopic reduction pipeline that uses Gaussian
processes or flexible model for the spectral extraction or wavelength
calibration, so we have chosen to leave the description in this paper. We placed
it in the appendix because we agree that, while cumbersome, does describe a
novel reduction procedure. Most readers can just ignore the appendix.

> Other suggested changes:
> ========================

> Section 1, paragraph 5:
> -----------------------
> "is found to be alpha~1.5-2 out to..." Since the alpha power law is determined
> for the distribution of separations, the sentence should be reversed here: "is
> found to be alpha~1.5-2 out to projected separations ~0.65pc (~0.25degrees of
> angular separation on the sky at a distance of 150pc),"

Fixed.

> Section 1, paragraph 7:
> -----------------------
> "that falls of until" -> "that falls off until"

Fixed.

> Section 2.1, paragraph 2:
> -------------------------
> "...likelihood ratio for all stars within 10pc and with a
> difference..."
>
> Perhaps you mean:
>
> "...likelihood ratio for a star to have another star within 10 pc of
> itself, and with a difference..."

For each star in TGAS, we compute the likelihood ratio for all nearby stars
(within 10pc). We've restructured this sentence to try to clarify.

> Section 2.1, paragraph 3:
> -------------------------
> "...exclusive star pairs within 200 pc in Heliocentric distance."
> ->
> "...exclusive stars pairs with heliocentric distances d<200pc.

Fixed.

> "...used above and given the limitations of the TGAS catalog itself..."
> ->
> "...used above, and within the limits of the TGAS catalog itself
> (e.g. magnitude range), TGAS is..."

Fixed.

> "...with airmass sec z < 1.5..."
> ->
> [The airmass is an incidental parameter of the observations, not an
> intrinsic property of the targets. This statement should therefore be
> moved to section 2.2.]

Fixed.

> "...all TGAS stars within 200 pc."
> ->
> [Please cite the actual number of TGAS stars within 200pc.]

Fixed. In fact, we fixed this to read "all 195,294 TGAS stars with 2MASS
photometry within 200 pc...".