### 2018-04-08 paper/submit/referee-response-2.txt
ref reports

We thank the referee for the suggestions. We hope we have addressed the major
concerns by re-doing the mixture modeling and adding text to the relevant
sections of this paper. The referee report is included below preceded by ">"
characters, with our responses below points.

> The statistical analysis presented in section 3 appears to be an
> attempt at modeling the 3D velocity differences (\psi) in a subset of
> 311 co-moving pairs, using a combination of a model that represents
> the distribution of velocity differences for real co-moving pairs,
> with one that represents the distribution of velocity differences for
> randomly selected field stars.

> While the model for the distribution of co-moving pairs (P_{co})
> appears satisfactory, the model for the distribution of randomly
> selected field stars (P_{field}) is not.

> Here's the problem:
> The subset of 311 co-moving candidates represents
> a subset of objects that have been *pre-selected* on the basis of
> having similar proper motions (from TGAS). As a result, their
> distribution of velocity differences cannot be represented by the 3D
> distribution of field stars, because these pairs represent a subset of
> objects that *a priori* have motions that are much more similar to
> each other than to normal field stars. Effectively, this subset
> represent stars that are already known to have motions that are
> similar to each other along two components of motion (motion in the
> plane of the sky), and thus the only uncertainty that remains is in
> the radial velocity direction - in other words, this is a 1D problem,
> and not a 3D problem. If you pre-select stars that have identical
> transverse motion, then the distribution of velocity differences is
> going to be a truncated Gaussian, and not a Maxwellian distribution.

> As a result, the modeled probability distribution P_{field} grossly
> underestimates the likelihood that two stars from the subset of common
> proper motion pairs will have total velocity differences near 0 km/s.

> Unfortunately, if that is what was done, the entire analysis in section 3,
> as well as the general conclusions of the manuscript, are invalid.

> Perhaps I have misunderstood the analysis presented in section
> 3. If that is the case, then I would urge the authors to present a
> revised discussion that would explain more clearly what has been done,
> in particular one that would clearly explain how the
> simulationaccounts for the fact that we are dealing with a subset of
> stars that were initially preselected based on having similar proper
> motions, and how this affects the results.

Our selection of the candidates does not involve selecting common proper motion
pairs of stars: as is described in this and in previous work, we ask whether two
stars are consistent with having the same 3D Cartesian velocity given parallax
and proper motion components. This is subtly different, however we do
acknowledge that this will skew the velocity-difference distribution of field
pairs, and the previous mixture modeling may be biased because of selection
effects.

As a more conservative and much simpler model for the comoving fraction, f, we
have re-done the analysis by modeling only the radial velocity differences
between pairs.

We think the referee has misunderstood the interpretation of "f" (the comoving
probability), so we have updated much of the language to try to clarify. f is
not the probability that a given pair is a coeval, comoving pair of stars, but
rather it is just the probability that two stars are comoving in three
dimensions. Given this distinction and the assumptions specified in the text, we
think our model is specified correctly. We have added the following text to the
section on modeling (see bold text in Section 3):

"We note that $f$ is \emph{not} the fraction of pairs that are coeval and
comoving; here we only model the fraction of pairs that are consistent with
being comoving in 3D velocity space given our measurements, some of which could
be chance alignments of stars in full-space position and velocity.
Therefore, $f$ is not a measure of our contamination rate due to random
alignments of comoving stars."

We have also added a similar note to the beginning of the results section:

"We again note that $f$ is not a measure of contamination from chance
field star pairings, but rather just quantifies the fraction of pairs within our
observed sample that are consistent with having a small 3D velocity difference."

In editing the text to clarify this point, we also have softened some of our
claims based on the resulting high-probability pairs. Without also modeling the
total distribution function of the Milky Way and all selection effects, we can't
actually know the contamination fraction of unassociated stars with the same 3D
velocity at large separations. From the LAMOST data visualized in this work, and
from recent work by Andrews et al., large separation (>1 pc) comoving in 3D
pairs with the same chemical abundances do exist, and so we now only state that
further follow-up of our sample of confidently 3D comoving pairs should allow
for a refined sample that contains likely coeval, very wide separation pairs. We
have added some text to the results section to try to clarify this - see bolded
text.

> As pointed out in the first report, it is surprising that so many of
> these common proper motion pairs (selected from Oh et al. 2017) are
> found after these follow-up spectroscopic observations to have
> velocity differences >20km/s, especially for the subset of pairs with
> separations 1pc-10pc. Remember that these are all common proper motion
> pairs, so their 3D motions must be relatively small to begin with;
> effectively their velocity differences start with being close to 0km/s
> when the radial velocity is not included, and then grows to pretty
> large values when the radial velocity data is included. This suggests
> a high level of contamination in the sample, from which one could
> argue that the majority of pairs with distances 1pc-10pc are in fact
> chance aalignment This is a serious concern that must be addressed.

We agree that there is likely significant contamination in the Oh et al. 2017
sample at large separations; this is stated in the original paper, and in this
paper text. One clarification: Oh et al. 2017 do not select pairs as being
"common proper motion," but rather construct a statistical model that accounts
for projection effects when comparing stars with the same 3D velocity that have
large (>~ 1 deg) separations.

> On the other hand, it is clear that of the ~20-25 pairs in the <1pc
> subset, the majority of them appear to be confirmed as co-moving
> objects. The manuscript should probably focus on these objects. For
> pairs with separations >1pc, the present study does not provide
> sufficient evidence to identify any of these objects as a probable
> co-moving system, and instead appear to suggest that most of them are
> chance aalignments This can only be solved with more accurate radial
> velocity information.

We believe that we have shown that some fraction of the large separation pairs
are consistent with being comoving, but agree that we can't conclude whether
these systems are coeval or chance alignments (and have stated this in the
text). We feel that it is still worth discussing these systems because of their
great interest. Precisely because contamination is an issue at large
separations, by identifying pairs that appear to be comoving with low-resolution
spectroscopy, we can improve yields for follow-up with higher resolution
spectroscopy (as is planned and being pursued by other members of our group).
We feel that we have stressed this point in the discussion, but have added some
qualifying statements to the conclusions section to help clarify.

> Other recommended corrections:
> ==============================

> Abstract:
> ---------

> "Of the 311 observed comoving pairs, we confirm 124 comoving pairs
> have probabilities 50% with separations as large as 10 pc, our
> original search limit."
> =>
> "Of an initial sample of 311 possible comoving pairs targeted for
> follow-up spectroscopy, we identify 124 pairs that have co-moving
> probabilities >50% with separations as large as 10 pc, our original
> search limit."

Changed.

> "...the number of high-probability pairs per separation decreases..."
> =>
> "...the number of high-probability pairs per separation unit
> decreases..."

Changed.

> "From ~0.5 pc to 10 pc, the number per separation is approximately
> uniform..."
> =>
> "From ~0.5 pc to 10 pc, on the other hand, the number per separation
> is approximately uniform..."

Changed.

> "We thus confirm the discovery of a population of comoving..."
> =>
> "This supports the existence of a population of comoving..."

Changed.

> Section 3:
> ----------

> "...is comoving in 3D velocity given..."
> =>
> "...is comoving in 3D velocity space given..."

Changed.